Global patterns of forest autotrophic carbon fluxes

### 2019-09-12 manuscript/manuscript draft.Rmd
underscore for CO~2~

Carbon fixation, allocation, and metabolism by trees sets the basis for energy and material flows in forest ecosystems and defines their interactions with Earth's changing climate. Forest autotrophic carbon flux (FACF) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**). Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**). On the global scale, forests play a critical role in regulating atmospheric CO_2 and climate (**Bonan**). The total amount of CO_2 cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**). 
Given the importance of understanding broad-scale patterns in forest autotrophic carbon flux (FACF), there have been many meta-analyses on the theme. However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time. The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple components of forest productivity and the principle climatic drivers of these patterns.

>>>>

Carbon fixation, allocation, and metabolism by trees sets the basis for energy and material flows in forest ecosystems and defines their interactions with Earth's changing climate. Forest autotrophic carbon flux (FACF) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**). Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**). On the global scale, forests play a critical role in regulating atmospheric CO~2~ and climate (**Bonan**). The total amount of CO~2~ cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**). 
Given the importance of understanding broad-scale patterns in forest autotrophic carbon flux (FACF), there have been many meta-analyses on the theme. However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time. The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple components of forest productivity and the principle climatic drivers of these patterns.
